---
title: "My (Everyday) R Notes"
author: "Congli (Claire) Zhang"
date: "6/28/2021"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require(pacman)) install.packages('pacman', repos = 'https://cran.rstudio.com')
pacman::p_load(tidyverse, dplyr, knitr, lfe, arsenal, ggpubr, broom, stargazer, fixest, modelsummary, kableExtra)
theme_set(theme_classic())
```

Organize the note using:
header 1, area of use, e.g., visualization, clean and tidy data, modeling, etc.
header 2, package
header 3, function
header 4, whatever important heading for that function

```{r}
head(mtcars)
```

# Importing exporting data sets
```{r}
here::i_am("code/filename.R")
fwrite(df, "data/df.csv")
```


# Cleaning and tidying data
## R basic
### table()
create tabular results of categorical variables.
check for one variable:
```{r}
table(mtcars$cyl) # does not report NA's
table(mtcars$cyl, exclude = NULL) # reports NA's
table(mtcars$cyl, exclude = "4")
table(mtcars$gear > 4)
table(mtcars$gear > 4, useNA = "always")
```

contingency table for two variables:
```{r}
table(mtcars$cyl, is.na(mtcars$carb))
```

## dplyr
### mutate()

recode:
```{r}
foo <- mtcars %>% 
  mutate(cyl = recode(cyl, "4" = "four", "6" = "six", "8" = "eight"))
```

panel data strategy - you may want to set flag for cases when you have missing values on a time-invariant variable, gender for example, you want to code the missing value from a certain year by using the non-missing value from other years:
```{r}
foo <- mtcars %>%
  group_by(cyl) %>% 
  arrange(gear) %>% 
  mutate(last_obs_fg = row_number() == n()) %>% 
  mutate(carb_new = ifelse(carb == 4, carb, carb[last_obs_fg == TRUE]))
```

