---
title: "My R Notes for DID Estimator"
author: "Congli (Claire) Zhang"
date: "Starting at: 8/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
if (!require(pacman)) install.packages('pacman', repos = 'https://cran.rstudio.com')
pacman::p_load(tidyverse, dplyr, knitr, lfe, arsenal, ggpubr, broom, stargazer, fixest, gtsummary, huxtable)
theme_set(theme_classic())
```


## An extraordinary example replicating the Card and Krueger (1994)  paper

The famous [Card and Krueger (1994) paper](https://www.nber.org/papers/w4509) examines the effects of minimum wages on employment. Philipp Leppert [replicates this study in R](https://rpubs.com/phle/r_tutorial_difference_in_differences).

### Run the code 

Dataset can be downloaded from  [David Card's page](https://davidcard.berkeley.edu/data_sets.html).

```{r, the raw data}
# Temporary file and path
tfile_path <- tempfile()
tdir_path <- tempdir()

# Download zip file
download.file("http://davidcard.berkeley.edu/data_sets/njmin.zip", 
              destfile = tfile_path)

# Unzip
unzip(tfile_path, exdir = tdir_path)

# Read codebook
codebook <- readr::read_lines(file = paste0(tdir_path, "/codebook"))

# Generate a vector with variable names
variable_names <- codebook %>%
  `[`(8:59) %>% # Variablennamen starten bei Element 8 (sheet)
  `[`(-c(5, 6, 13, 14, 32, 33)) %>% # Elemente ohne Variablennamen entfernen
  str_sub(1, 8) %>% # längster Variablenname enthält 8 Zeichen
  str_squish() %>% # Whitespaces entfernen
  str_to_lower() # nur Kleinbuchstaben verwenden

# Generate a vector with variable labels
variable_labels <- codebook %>%
  `[`(8:59) %>% # variable names start at element 8 (sheet)
  `[`(-c(5, 6, 13, 14, 32, 33)) %>% # remove elements w/o variable names
  sub(".*\\.[0-9]", "", .) %>%
  `[`(-c(5:10))  %>% # these elements are combined later on
  str_squish() # remove white spaces
  
# Region
variable_labels[41] <- "region of restaurant"

# Read raw data
data_raw <- read_table2(paste0(tdir_path, "/public.dat"),
                        col_names = FALSE)
```


```{r, cleaned data}
# Add variable names
data_mod <- data_raw %>%
  select(-X47) %>% # remove empty column
  `colnames<-`(., variable_names) %>% # Assign variable names
  mutate_all(as.numeric) %>% # treat all variables as numeric
  mutate(sheet = ifelse(sheet == 407 & chain == 4, 408, sheet)) # duplicated sheet id 407

# Process data (currently wide format)
data_mod <- data_mod %>%
  # chain value label
  mutate(chain = case_when(chain == 1 ~ "bk",
                           chain == 2 ~ "kfc",
                           chain == 3 ~ "roys",
                           chain == 4 ~ "wendys")) %>%
  # state value label
  mutate(state = case_when(state == 1 ~ "New Jersey",
                           state == 0 ~ "Pennsylvania")) %>%
  # Region dummy
  mutate(region = case_when(southj == 1 ~ "southj",
                            centralj == 1 ~ "centralj",
                            northj == 1 ~ "northj",
                            shore == 1 ~ "shorej",
                            pa1 == 1 ~ "phillypa",
                            pa2 == 1 ~ "eastonpa")) %>%
  # meals value label
  mutate(meals = case_when(meals == 0 ~ "none",
                           meals == 1 ~ "free meals",
                           meals == 2 ~ "reduced price meals",
                           meals == 3 ~ "both free and reduced price meals")) %>%
  # meals value label
  mutate(meals2 = case_when(meals2 == 0 ~ "none",
                            meals2 == 1 ~ "free meals",
                            meals2 == 2 ~ "reduced price meals",
                            meals2 == 3 ~ "both free and reduced price meals")) %>%
  # status2 value label
  mutate(status2 = case_when(status2 == 0 ~ "refused second interview",
                             status2 == 1 ~ "answered 2nd interview",
                             status2 == 2 ~ "closed for renovations",
                             status2 == 3 ~ "closed permanently",
                             status2 == 4 ~ "closed for highway construction",
                             status2 == 5 ~ "closed due to Mall fire")) %>%
  mutate(co_owned = if_else(co_owned == 1, "yes", "no")) %>%
  mutate(bonus = if_else(bonus == 1, "yes", "no")) %>%
  mutate(special2 = if_else(special2 == 1, "yes", "no")) %>%
  mutate(type2 = if_else(type2 == 1, "phone", "personal")) %>%
  select(-southj, -centralj, -northj, -shore, -pa1, -pa2) %>% # now included in region dummy
  mutate(date2 = lubridate::mdy(date2)) %>% # Convert date
  rename(open2 = open2r) %>% #Fit name to wave 1
  rename(firstinc2 = firstin2) %>% # Fit name to wave 1
  sjlabelled::set_label(variable_labels) # Add stored variable labels
```

The data is in a wide format. Need to tidy it.

```{r, transposed data}
# Structural variables
structure <- data_mod %>%
  select(sheet, chain, co_owned, state, region)

# Wave 1 variables
wave1 <- data_mod %>%
  select(-ends_with("2"), - names(structure)) %>%
  mutate(observation = "February 1992") %>%
  bind_cols(structure) 

# Wave 2 variables
wave2 <- data_mod %>%
  select(ends_with("2")) %>%
  rename_all(~str_remove(., "2"))  %>%
  mutate(observation = "November 1992") %>%
  bind_cols(structure) 

# Final dataset
card_krueger_1994 <- bind_rows(wave1, wave2) %>%
  select(sort(names(.))) %>% # Sort columns alphabetically
  sjlabelled::copy_labels(data_mod) # Restore variable labels
```

According to the paper, the full-time equivalents (FTE) consist of full-time employees, managers and part-time employees (emptot). The latter are multiplied by factor 0.5 before entering the calculation. Also, generate the share of full-time employees of all FTE (pct_ftw).

```{r, final data}
card_krueger_1994_mod <- card_krueger_1994 %>%
  mutate(emptot = empft + nmgrs + 0.5 * emppt,
         pct_fte = empft / emptot * 100)
```

Some descriptive stats. Before get to Leppert's code, I think tbl_strata() does a pretty neat job here, except for one thing, the table displays median and IQR for continuous variables and the changing statistics code doesn't seem to work - NEED FIX THIS LATER.

```{r}
tbl_strata(card_krueger_1994_mod,
           strata = observation,
           .tbl_fun =
             ~ .x %>%
             tbl_summary(by = state, missing = "no") %>%
             add_n(),
           statistic = list(all_continuous() ~ "{mean} ({sd})",
                            all_categorical() ~ "{n} / {N} ({p}%)"))
```


```{r, relative shares of restaurants within each state}
card_krueger_1994_mod %>%
  select(chain, state) %>%
  table() %>%
  prop.table(margin = 2)  %>%
  apply(MARGIN = 2,
        FUN = scales::percent_format(accuracy = 0.1)) %>%
  noquote
```


```{r, pre-treatment means}
card_krueger_1994_mod %>%
  filter(observation == "February 1992") %>%
  group_by(state) %>%
  summarise(emptot = mean(emptot, na.rm = TRUE),
            pct_fte  = mean(pct_fte, na.rm = TRUE),
            wage_st = mean(wage_st, na.rm = TRUE),
            hrsopen = mean(hrsopen, na.rm = TRUE)) %>%
  pivot_longer(cols=-state, names_to = "variable") %>%
  pivot_wider(names_from = state, values_from = value)
```


```{r, post-treatment means}
card_krueger_1994_mod %>%
  filter(observation == "November 1992") %>%
  group_by(state) %>%
  summarise(emptot = mean(emptot, na.rm = TRUE),
            pct_fte  = mean(pct_fte, na.rm = TRUE),
            wage_st = mean(wage_st, na.rm = TRUE),
            hrsopen = mean(hrsopen, na.rm = TRUE)) %>%
  pivot_longer(cols=-state, names_to = "variable") %>%
  pivot_wider(names_from = state, values_from = value)
```

Figure 1 in the study; The authors created this figure with SAS and it shows the distribution of the fast food restaurants’ wages grouped by the federal states NJ and PA before and after the treatment.

```{r}
hist.feb <- card_krueger_1994_mod %>%
  filter(observation == "February 1992") %>%
  ggplot(aes(wage_st, fill = state)) +
  geom_histogram(aes(y=c(..count..[..group..==1]/sum(..count..[..group..==1]),
                         ..count..[..group..==2]/sum(..count..[..group..==2]))*100),
                 alpha=0.5, position = "dodge", bins = 23) +
  labs(title = "February 1992", x = "Wage range", y = "Percent of stores", fill = "") +
  scale_fill_grey()

hist.nov <- card_krueger_1994_mod %>%
  filter(observation == "November 1992") %>%
  ggplot(aes(wage_st, fill = state)) +
  geom_histogram(aes(y=c(..count..[..group..==1]/sum(..count..[..group..==1]),
                         ..count..[..group..==2]/sum(..count..[..group..==2]))*100),
                 alpha = 0.5, position = "dodge", bins = 23) +
  labs(title = "November 1992", x="Wage range", y = "Percent of stores", fill="") +
  scale_fill_grey()

ggpubr::ggarrange(hist.feb, hist.nov, ncol = 2, 
                  common.legend = TRUE, legend = "bottom")
```

Visualizing DID

 * first differences; obtain four means: treat-before, control-before, treat-after, control-after

```{r, first difference}
differences <- card_krueger_1994_mod %>%
  group_by(observation, state) %>%
  summarise(emptot = mean(emptot, na.rm = TRUE))

# Treatment group (NJ) before treatment
njfeb <- differences[1,3]

# Control group (PA) before treatment
pafeb <- differences[2,3]

# Treatment group (NJ) after treatment
njnov <- differences[3,3]

# Control group (PA) after treatment
panov <- differences[4,3]
```

 * ATT, which is the difference between the difference of November and February within NJ and PA. two results should be identical

```{r, the average treatment effect - ATT}
(njnov-njfeb)-(panov-pafeb) 

(njnov-panov)-(njfeb-pafeb)
```

 * visualizing ATT
 
```{r}
# Calculate counterfactual outcome
nj_counterfactual <- tibble(
  observation = c("February 1992","November 1992"), 
  state = c("New Jersey (Counterfactual)","New Jersey (Counterfactual)"),
  emptot = as.numeric(c(njfeb, njfeb-(pafeb-panov)))
  ) 

# Data points for treatment event
intervention <- tibble(
    observation = c("Intervention", "Intervention", "Intervention"),
    state = c("New Jersey", "Pennsylvania", "New Jersey (Counterfactual)"),
    emptot = c(19.35, 22.3, 19.35)
  ) 

# Combine data
did_plotdata <- bind_rows(differences, 
                          nj_counterfactual, 
                          intervention)
did_plotdata %>%
  mutate(label = if_else(observation == "November 1992", as.character(state), NA_character_)) %>%
  ggplot(aes(x=observation,y=emptot, group=state)) +
  geom_line(aes(color=state), size=1.2) +
  geom_vline(xintercept = "Intervention", linetype="dotted", 
             color = "black", size=1.1) + 
  scale_color_brewer(palette = "Accent") +
  scale_y_continuous(limits = c(17,24)) +
  ggrepel::geom_label_repel(aes(label = label),
                   nudge_x = 0.5, nudge_y = -0.5,
                   na.rm = TRUE) +
  guides(color=FALSE) +
  labs(x="", y="FTE Employment (mean)") +
  annotate(
    "text",
    x = "November 1992",
    y = 19.6,
    label = "{Difference-in-Differences}",
    angle = 90,
    size = 3
  )
```

Finally, get to the DID estimator,

```{r}
card_krueger_1994_mod <- mutate(card_krueger_1994_mod,
                                time = ifelse(observation == "November 1992", 1, 0),
                                treated = ifelse(state == "New Jersey", 1, 0)
                                )

lm(emptot ~ time + treated + time:treated, data = card_krueger_1994_mod) %>% 
  huxtable::huxreg()
```

To more precisely estimate, add sheet fixed-effects, for some reason, Leppert's code doesn't work for me. I'll go ahead using fixest package to obtain same results.

```{r}
# # Declare as panel data
# panel <- pdata.frame(card_krueger_1994_mod, "sheet")
# 
# # Within model
# did.reg <- plm::plm(emptot ~ time + treated + time:treated,
#                     data = panel, 
#                     model = "within")
# 
# # obtain clustered standard errors
# lmtest::coeftest(did.reg, vcov = function(x) 
#   vcovHC(x, cluster = "group", type = "HC1"))
```


```{r}
feols(emptot ~ time + treated + time:treated | sheet,
      card_krueger_1994_mod) %>% 
  huxreg()
```



