---
title: "My R Notes for Data management"
author: "Congli (Claire) Zhang"
date: "Starting at: 6/28/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
if (!require(pacman)) install.packages('pacman', repos = 'https://cran.rstudio.com')
pacman::p_load(tidyverse, dplyr, knitr, lfe, arsenal, ggpubr, broom, stargazer, fixest, modelsummary, kableExtra)
theme_set(theme_classic())
```

This is my notes of R code for data manipulation and management.

Organizing the notes using:

 * header 2, area of use, e.g., visualization, clean and tidy data, modeling, etc.
 * header 3, package::function
 * header 4, whatever important heading for that function


## 1. Importing and exporting data

When you don't have a project created beforehand:
```{r}
# here::i_am("your code folder/your R or RMD file")
# datatable <- rio::import(here::here("your data folder", "your data file"))
# rio::export(your datatable, here::here("your data folder", "your data file"))
```

When you have an existing project:
```{r}
# datatable <- rio::import(here::here("your data folder", "your data file"))
# rio::export(your datatable, here::here("your data folder", "your data file"))
```


## 2. Inspecting the dataset

### Base R - handy, quick, and flexible

#### head()

arrange by a variable and show whatever number of observations you like:
```{r}
head(mtcars %>% arrange(desc(gear)), 10) 
```


#### table()

create tabular results of categorical variables. check for one variable:
```{r}
table(mtcars$cyl) # does not report NA's
table(mtcars$cyl, exclude = NULL) # reports NA's
table(mtcars$cyl, exclude = "4")
table(mtcars$gear > 4)
table(mtcars$gear > 4, useNA = "always")
```

contingency table for two variables:
```{r}
table(mtcars$cyl, is.na(mtcars$carb))
```

binned frequency table:
```{r}
table(cut(mtcars$cyl, seq(min(mtcars$cyl), max(mtcars$cyl), by = 0.5)))
```

### gtsummary::tbl_summary()

Instantly fell in love with this package/function when I was exhausting options to create the infamous "Table 1". Made a habit of doing this for first inspection of new data.

the default result:
```{r}
library(gtsummary)
tbl_summary(mtcars)
```

the default result for comparison table:
```{r}
tbl_summary(mtcars, by = cyl)
```

it's arguments are not super handy though:
```{r}
mtcars %>% 
  tbl_summary(by = cyl) %>%
  modify_header(label ~ "**Key Variables**")
```

just include a few variables:
```{r}
mtcars %>% 
  tbl_summary(by = cyl,
              include = c(cyl, mpg, vs, gear),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} / {N} ({p}%)")) %>% 
  add_p() %>% 
  modify_header(label ~ "**Key Variables**")
```


get fancier with the names of the variables and levels
```{r}
mtcars$gear <- factor(mtcars$gear, levels = c(3, 4, 5), labels = c("Three", "Four", "Five"))

mtcars %>% 
  tbl_summary(by = cyl,
              include = c(cyl, mpg, vs, gear),
              label = list(mpg = "MPG",
                           vs = "VS",
                           gear = "GEAR"),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} / {N} ({p}%)")) %>% 
  add_p() %>% 
  modify_header(label ~ "**Key Variables**")
```

## 3. Cleaning and tidying data

### dplyr::mutate()

recode:
```{r}
foo <- mtcars %>% 
  mutate(cyl = recode(cyl, "4" = "four", "6" = "six", "8" = "eight"))
```

panel data strategy - you may want to set flag for cases when you have missing values on a time-invariant variable, gender for example, you want to code the missing value from a certain year by using the non-missing value from other years:
```{r}
foo <- mtcars %>%
  group_by(cyl) %>% 
  arrange(gear) %>% 
  mutate(last_obs_fg = row_number() == n()) %>% 
  mutate(carb_new = ifelse(carb == 4, carb, carb[last_obs_fg == TRUE]))
```

